---
title: "04_describe"
format: html
editor: visual
---

# Clear Workspace
```{r}
rm(list = ls())
```

# Load libraries
```{r}
library("tidyverse")
library("ggthemes")
library("patchwork")
```

# Load data
```{r}
#| message: false
diabetes_data <- read_tsv(file = '../data/02_diabetes_data_clean.tsv',
                      na = 'NA')

```

# Observations and variables in the dataset

There are 6795 observation distributed in 20 different variables
```{r}
diabetes_data |>
  dim()
```

The variables have the different names
```{r}
diabetes_data |> 
  colnames()
```

Distribution of diagnosis
```{r}
non_diagnosed <- diabetes_data |> 
  filter(diagnosis == 0) |>
  count()

diagnosed <- diabetes_data |> 
  filter(diagnosis == 1) |>
  count()

medicated <- diabetes_data |> 
  filter(medication == 1) |>
  count()

table <- data.frame( non_diagnosed, diagnosed, medicated)
rownames(table) <- 'Count'
colnames(table) <- c('Non diagnosed', 'Diagnosed', 'Medicated')
print(table)


```


# Gender distribution

The gender distribution in the study population is even distributed with 3423 female and 3372 male

```{r}
gender_count <- diabetes_data |>
  group_by(sex) |>
  count() 
gender_count
```

# Age distribution plot

```{r}
diabetes_data |> 
  filter(sex == 'female' |
         sex == 'male') |>

    mutate(AgeGroup = case_when(
      age > 10 & age <= 19 ~ '(10,20]',
      age > 20 & age <= 29 ~ '(20,30]',
      age > 30 & age <= 39 ~ '(30,40]',
      age > 40 & age <= 49 ~ '(40,50]',
      age > 50 & age <= 59 ~ '(50,60]',
      age > 60 & age <= 69 ~ '(60,70]',
      age > 70 & age <= 80 ~ '(70,80]')) |>
      
   mutate(AgeGroup = factor(AgeGroup,
                        levels = c('(10,20]',
                                   '(20,30]',
                                   '(30,40]',
                                   '(40,50]',
                                   '(50,60]',
                                   '(60,70]',
                                   '(70,80]'))) |>
  drop_na() |>

  ggplot(mapping = aes(x = AgeGroup,
                       fill = sex)) +
  geom_bar(position='dodge',
           alpha = 0.9,
           color='black') + 
  scale_color_colorblind() +
  theme_minimal() +
   labs(title = 'Age distribution of the study population',
       x = 'Age groups',
       y = 'Number of people') +
  theme(legend.position = 'bottom')
```

A variable to consider here is the age, as the occurrence of observations with the age = 80 is around double as high as the next highest occurrence. 
 
```{r}
age_dist <- ggplot(data = diabetes_data,
                   mapping = aes(x= age)) +
geom_histogram(color='blue',
               fill='light blue',
               binwidth = 1)
age_dist
```
This might be because this value is set to be 80+ in the study, and therefore includes everyone that is 80 years or older. 



# BMI distribution of the study population

```{r}
diabetes_data |>
  ggplot(aes(x = bmi)) +
  annotate("rect", fill = "green", alpha = 0.4, 
        xmin=18.5, xmax=24.9, ymin=0, ymax=+Inf) +
  geom_histogram(color='darkblue', fill='lightblue', binwidth = 2) +
  theme_minimal() +
  labs(title = 'BMI distribution in the study population',
       x = bquote('BMI'~(kg/m^2)),
       y = 'Number of people')
```
We see a slightly overweight BMI average for the population of the study.

The BMI distribution for the study population based on the diabetes diagnose
for non-medicated people

```{r}
non_medicated_data <- diabetes_data |>
  filter(medication == 0)
ggplot(non_medicated_data,
       mapping = aes(x = factor(diagnosis),
                     y = bmi)) +
geom_boxplot(aes(colour = factor(diagnosis))) +
  theme_minimal() +
  labs(title = 'BMI distribution in the study population',
       x = 'Diabetes diagnosed',
       y = bquote('BMI'~(kg/m^2)),
       color = 'Diagnosis') +
  theme(legend.position = 'bottom')
  
```
We see from the plot above that people diagnosed with DM or pre-DM generally have a slightly higher BMI average.

# BMI compared based on ages intervals
```{r}
non_medicated_data |> 
  mutate(AgeGroup = case_when(
      age > 10 & age <= 19 ~ '(10,20]',
      age > 20 & age <= 29 ~ '(20,30]',
      age > 30 & age <= 39 ~ '(30,40]',
      age > 40 & age <= 49 ~ '(40,50]',
      age > 50 & age <= 59 ~ '(50,60]',
      age > 60 & age <= 69 ~ '(60,70]',
      age > 70 & age <= 80 ~ '(70,80]')) |>
      
   mutate(AgeGroup = factor(AgeGroup,
                        levels = c('(10,20]',
                                   '(20,30]',
                                   '(30,40]',
                                   '(40,50]',
                                   '(50,60]',
                                   '(60,70]',
                                   '(70,80]'))) |>
  drop_na() |>

  ggplot(mapping = aes(x = AgeGroup,
                       y = bmi,
                       fill = factor(diagnosis))) +
  geom_boxplot(aes(color = factor(diagnosis)),
           alpha = 0.9,
           color='black') + 
  scale_color_colorblind() +
  theme_minimal() +
   labs(fill = 'Diagnosis',
        title = 'BMI compared to age',
        x = 'Age groups',
        y = 'BMI') +
  theme(legend.position = 'bottom')
```

The data includes 5 different ethnicity groups with the following observations from each group

```{r}
f1 <- diabetes_data |> 
  filter(ethnicity == 'Mexican American') |>
  count()

f2 <- diabetes_data |> 
  filter(ethnicity == 'Non-Hispanic White') |>
  count()

f3 <- diabetes_data |> 
  filter(ethnicity == 'Non-Hispanic Black') |>
  count()

f4 <- diabetes_data |> 
  filter(ethnicity == 'Other Hispanic') |>
  count()

f5 <- diabetes_data |> 
  filter(ethnicity == 'Other Race Including Multi-Racial') |>
  count()

Ethnicities <- data.frame(f1, f2, f3, f4, f5)
rownames(Ethnicities) <- 'Count'
colnames(Ethnicities) <- c('Mexican American', 'Non-Hispanic White', 'Non-Hispanic Black','Other Hispanic','Other Race Including Multi-Racial')
print(Ethnicities)
```

# Diagnosis and Medication grouped by Ethnicity
```{r}
combined_plot <- ggplot(diabetes_data, 
                        aes(x = ethnicity, 
                            fill = factor(medication), 
                            group = interaction(ethnicity, factor(diagnosis)))) +
  geom_bar(aes(y = ..count..), 
           position = position_dodge(width = 0.7), 
           stat = "count") +
  scale_fill_manual(values = c("red", "blue"), 
                    labels = c("Not Medicated", "Medicated")) +
  labs(title = "Diagnosis and Medication by Ethnicity",
       fill = "Medicated",
       x = "Ethnicity",
       y = "Count") +
  facet_wrap(~ factor(diagnosis), scales = "free_x", 
             labeller = as_labeller(c('0' = "Non Diagnosed", '1' = "Diagnosed"))) +
  theme_minimal()

print(combined_plot)

```

