---
title: "04_describe"
format:
  html:
    embed-resources: true
editor: visual
---

# Clear Environment

```{r}
rm(list = ls())
```

# Load libraries

```{r}
#| message: false
library("tidyverse")
library("ggthemes")
```

# Load data

```{r}
#| message: false
diabetes_data <- read_tsv(file = '../data/03_diabetes_data_augment.tsv',
                          na = 'NA')

```

# Observations and variables in the dataset

There are 6795 observation distributed in 20 different variables

```{r}
diabetes_data |>
  dim()
```

The variables have the different names

```{r}
diabetes_data |> 
  colnames()
```

Distribution of diagnosis

```{r}
diabetes_data |>
  count(dm_status)
```

# Gender distribution

The gender distribution in the study population is evenly distributed with 3423 females and 3372 males.

```{r}
diabetes_data |>
  count(sex)
```

# Age distribution plot

A variable to consider here is the age, as the occurrence of observations with the age = 80 is around double as high as the next highest occurrence.

```{r}
age_dist <- diabetes_data |> 
  ggplot(mapping = aes(x= age)) +
  geom_histogram(color='blue',
               fill='light blue',
               binwidth = 1) +
  labs(title = 'Age distribution of the study population',
       x = 'Age',
       y = 'Number of people') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# save plot to results 
ggsave(age_dist, 
       filename = '../results/04_age_dist.png',
       device = "png",
       height = 10, 
       width = 20, 
       units = "cm")

age_dist

```

This might be because this value is set to be 80+ in the study, and therefore includes everyone that is 80 years or older.

# BMI distribution of the study population

```{r}
bmi_dist <- diabetes_data |>
  ggplot(aes(x = bmi)) +
  annotate("rect", 
           fill = "green", 
           alpha = 0.4, 
           xmin=18.5, 
           xmax=24.9, 
           ymin=0, 
           ymax=+Inf) +
  geom_histogram(color='darkblue', 
                 fill='lightblue', 
                 binwidth = 2) +
  theme_minimal() +
  labs(title = 'BMI distribution in the study population',
       x = bquote('BMI'~(kg/m^2)),
       y = 'Number of people') +
  theme(plot.title = element_text(hjust = 0.5))
  
# save plot to results 
ggsave(bmi_dist, 
       filename = '../results/04_bmi_dist.png',
       device = "png",
       height = 10, 
       width = 20, 
       units = "cm")

bmi_dist
```

We see a slightly overweight BMI average for the population of the study as the green area marks the normal BMI range.

The BMI distribution for the study population based on the diabetes diagnosis for non-medicated people

```{r}
non_medicated_data <- diabetes_data |>
  filter(medication == 0) |> 
  mutate(diagnosis = factor(diagnosis))

bmi_dist_diagnosed <- non_medicated_data |> 
  ggplot(mapping = aes(x = diagnosis,
                       y = bmi)) +
  geom_boxplot(aes(colour = diagnosis)) +
  theme_minimal() +
  scale_color_colorblind() +
  labs(title = 'BMI distribution in the study population',
       x = 'Diabetes diagnosis',
       y = bquote('BMI'~(kg/m^2)),
       color = 'Diagnosis') +
  theme(legend.position = 'bottom') +
  theme(plot.title = element_text(hjust = 0.5))
  
# save plot to results 
ggsave(bmi_dist_diagnosed, 
       filename = '../results/04_bmi_dist_diagnosed.png',
       device = "png",
       height = 10, 
       width = 20, 
       units = "cm")

bmi_dist_diagnosed
```

We see from the plot above that people diagnosed with DM or pre-DM generally have a slightly higher BMI average.
