---
title: "Logistic Regression Model"
format:
  html:
    embed-resources: true
editor: visual
---

# Clear Environment

```{r}
rm(list = ls())
```

# Load Libraries

```{r}
#| message: false
library("tidyverse")
library("patchwork")
library("MASS")
```

# Load data

```{r}
#| message: false
diabetes_data <- read_tsv(file = "../data/03_diabetes_data_augment.tsv",
                          na = "NA")
```

# Logistic regression model

Logistic regression modelling is performed to predict the variables that gives increase or decrease in the probability of a DM/pre-DM diagnosis. The model is based on anthropometric measurements and biomarkers.

# Backwards selection model

Backwards selection to find to most significant variables

```{r}
# Riskfactor for the logistic regression model
RiskFactor <- diabetes_data |>
  filter(medication == 0) |>
  dplyr::select(diagnosis,
                weight,
                height,
                leg,
                arml,
                armc,
                waist,
                tri_skinfold,
                sub_skinfold,
                gh,
                albumin,
                bun,
                creatinine)
```

```{r}
#Replace the NA with 0 in the riskfactors
RiskFactor <- RiskFactor |>
  drop_na()

```

```{r}
# Performing backwards selection to find the best parameters for the logistic regression model
model_full <- glm(diagnosis~.,
                  family = "binomial",
                  data = RiskFactor)

model_min <- glm(diagnosis ~ weight,
                 family = "binomial",
                 data = RiskFactor)


backwardsmodel <- stepAIC(model_full,
                          direction = 'backward',
                          scope = list(upper = model_full, 
                                       lower = model_min))
  

summarybackward <- tidy(backwardsmodel)

```

```{r}
# Add the odds ratio for the each variable to the summarybackward
summarybackward <- summarybackward |>
  mutate(OR = exp(estimate)) |>
  relocate(OR, .after = term)


```

```{r}
# Show the predicted variables found by the model
predictor <- summarybackward |> 
  dplyr::select(term) |>
  slice(2:n()) |>
  pull(term)


```

# Logistic regression model of the variables found from the backwards selection

```{r}
model_data <- diabetes_data |> 
  filter(medication == 0) |> 
  drop_na() |>
  dplyr::select(diagnosis,tidyselect::any_of(predictor)) |> 
  mutate(diagnosis = factor(diagnosis))

```

# Performing the logistic regression model

```{r}
model <- glm(diagnosis ~. ,
             family = "binomial",
             data = model_data)

summary(model)
```

# Plot the most significant variable

The logistics regression model for glycohemoglobin levels (Highest significant) and diabetes is plotted.

```{r}
LR_model <- diabetes_data |>
   ggplot(mapping = aes(gh, diagnosis)) +
   geom_point(alpha = 0.2) +
   geom_smooth(method = "glm",
               method.args = list(family = "binomial")) +
   labs(title = "Logistic Regression Model",
        x = "Glycohemoglobin concentration (%)",
        y = "Probability of diabetes") +
   theme_minimal() +
   theme(plot.title = element_text(hjust = 0.5)) 

LR_model
```

# Save the plots in results folder

```{r}
ggsave(LR_model, 
       filename = "../results/08_LR_model.png",
       device = "png",
       height = 15, width = 24, units = "cm")
```
